### 2020-08-31 code/figure4.Rmd
Modify Figure 4 to generate 2 panel figure (combining enriched / depleted into a single forest plot + heatmap)

up_p <- ggplot(up_compare_plot, 
               aes(x = or,
                   y = Hugo_Symbol,
                   color = cluster,
                   group = cluster)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_pointrange(aes(xmin = ci.low, xmax = ci.up)) +
  scale_color_manual(values = cluster_colors) +
  theme_classic() +
  xlab("log odds ratio") +
  xlim(c(0, 9)) +
  ylab("Gene") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y", switch = "y") +
  ggtitle("Enriched")
  theme(
    panel.grid.major.y = element_line(colour = "#f0f0f0"),
    axis.title.y = element_blank()
  )

dn_p <- ggplot(down_compare_plot, 
               aes(x = or, 
                   y = Hugo_Symbol, 
                   color = cluster, 
                   group = cluster)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_pointrange(aes(xmin = ci.low, xmax = ci.up)) +
  scale_color_manual(values = cluster_colors, guide = guide_none()) +
  theme_classic() +
  xlab("log odds ratio") +
  xlim(c(-9, 0)) +
  ylab("Gene") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y", switch = "y") +
  ggtitle("Depleted") +
  theme(panel.grid.major.y = element_line(colour = "#f0f0f0"))

tdat_up <- all_muts %>%
  filter(Hugo_Symbol %in% up_compare_plot$Hugo_Symbol) %>%
  distinct(Tumor_Sample_Barcode, Hugo_Symbol, clust_knn_char, type) %>%
  select(type, clust_knn_char, Hugo_Symbol) %>%
  table()
tdat_up <- left_join(
  as.data.frame(tdat_up),
  clinical_plus_cluster %>%
    select(Tumor_Sample_Barcode, type, clust_knn_char) %>%
    distinct() %>%
    group_by(type, clust_knn_char) %>%
    summarise(ntumors = n())) %>%
  mutate(prop = Freq / ntumors) %>%
  mutate(prop = ifelse(is.na(prop), 0, prop)) %>%
  select(type, clust_knn_char, Hugo_Symbol, prop) %>%
  rename(cluster = clust_knn_char)
tdat_up$Hugo_Symbol <- factor(tdat_up$Hugo_Symbol, levels = levels(up_compare_plot$Hugo_Symbol))
tdat_up <- up_compare_plot %>% left_join(tdat_up)

tdat_up_p <- tdat_up %>% ggplot(aes(type, Hugo_Symbol, fill = prop)) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", "Proportion of \nTumor Type \nMutated Within Cluster", limits = c(0, 1)) +
  xlab("Tumor Type") +
  ylab("Gene Symbol") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(colour = "#f0f0f0")
  )

tdat_dn <- all_muts %>%
  filter(Hugo_Symbol %in% down_compare_plot$Hugo_Symbol) %>%
  distinct(Tumor_Sample_Barcode, Hugo_Symbol, clust_knn_char, type) %>%
  select(type, clust_knn_char, Hugo_Symbol) %>%
  table()
tdat_dn <- left_join(
  as.data.frame(tdat_dn),
  clinical_plus_cluster %>%
    select(Tumor_Sample_Barcode, type, clust_knn_char) %>%
    distinct() %>%
    group_by(type, clust_knn_char) %>%
    summarise(ntumors = n())) %>%
  mutate(prop = Freq / ntumors) %>%
  mutate(prop = ifelse(is.na(prop), 0, prop)) %>%
  select(type, clust_knn_char, Hugo_Symbol, prop) %>%
  rename(cluster = clust_knn_char)
tdat_dn$Hugo_Symbol <- factor(tdat_dn$Hugo_Symbol, levels = levels(down_compare_plot$Hugo_Symbol))
tdat_dn <- down_compare_plot %>% left_join(tdat_dn)

tdat_dn_p <- tdat_dn %>% ggplot(aes(type, Hugo_Symbol, fill = prop)) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", "Proportion of \nTumor Type \nMutated Within Cluster", limits = c(0, 1)) +
  xlab("Tumor Type") +
  ylab("Gene Symbol") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(colour = "#f0f0f0")
  )

layout <- "ABBCDD"

dn_p + tdat_dn_p + up_p + tdat_up_p + plot_layout(guides = 'collect', design = layout) & theme(legend.position = "bottom")
```

>>>>

all_compare_plot <- bind_rows(down_compare_plot, up_compare_plot)
all_compare_plot %<>% arrange(or, .by_group = TRUE) %>% mutate(Hugo_Symbol = as.character(Hugo_Symbol)) %>% ungroup() %>% mutate(Row_Index = sprintf("%3i", row_number()))
Hugo_rownames <- all_compare_plot$Hugo_Symbol
names(Hugo_rownames) <- as.character(all_compare_plot$Row_Index)
all_p <- ggplot(all_compare_plot, 
               aes(x = or,
                   y = Row_Index,down_compare_plot <- all_compare %>%
  group_by(cluster) %>%
  filter(or < 0) %>%
  filter(adjPval < 0.01) %>%
  arrange(-or) %>%
  filter(is.finite(or)) %>%
  slice_max(order_by = or, n = 10) %>%
  arrange(cluster, or, .by.group = TRUE)
down_compare_plot$Hugo_Symbol <- with(down_compare_plot,
                                      factor(as.character(Hugo_Symbol),
                                             levels = unique(as.character(Hugo_Symbol))))

all_compare_plot <- bind_rows(down_compare_plot, up_compare_plot)
all_compare_plot %<>% arrange(or, .by_group = TRUE) %>% mutate(Hugo_Symbol = as.character(Hugo_Symbol)) %>% ungroup() %>% mutate(Row_Index = sprintf("%3i", row_number()))
Hugo_rownames <- all_compare_plot$Hugo_Symbol
names(Hugo_rownames) <- as.character(all_compare_plot$Row_Index)
all_p <- ggplot(all_compare_plot, 
               aes(x = or,
                   y = Row_Index,
                   color = cluster,
                   group = cluster)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_pointrange(aes(xmin = ci.low, xmax = ci.up)) +
  scale_y_discrete(labels = Hugo_rownames) +
  scale_color_manual(values = cluster_colors) +
  theme_classic() +
  xlab("log odds ratio") +
  ylab("Gene") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y", switch = "y") +
  theme(
    panel.grid.major.y = element_line(colour = "#f0f0f0"),
    axis.title.y = element_blank()
  )

tdat_all <- all_muts %>%
  filter(Hugo_Symbol %in% all_compare_plot$Hugo_Symbol) %>%
  distinct(Tumor_Sample_Barcode, Hugo_Symbol, clust_knn_char, type) %>%
  select(type, clust_knn_char, Hugo_Symbol) %>%
  table()
tdat_all <- left_join(
  as.data.frame(tdat_all),
  clinical_plus_cluster %>%
    select(Tumor_Sample_Barcode, type, clust_knn_char) %>%
    distinct() %>%
    group_by(type, clust_knn_char) %>%
    summarise(ntumors = n())) %>%
  group_by(type, clust_knn_char) %>%
  mutate(c_prop = Freq / ntumors) %>%
  mutate(c_prop = ifelse(is.na(c_prop), 0, c_prop)) %>%
  select(type, clust_knn_char, Hugo_Symbol, c_prop) %>%
  rename(cluster = clust_knn_char)
tdat_all <- all_compare_plot %>% left_join(tdat_all)
tdat_all_p <- tdat_all %>% ggplot(aes(type, Row_Index, fill = c_prop)) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", "Proportion of \nTumor Type \nMutated Within Cluster", limits = c(0, 1)) +
  xlab("Tumor Type") +
  ylab("Gene Symbol") +
  scale_y_discrete(labels = Hugo_rownames) +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(colour = "#f0f0f0")
  )

layout <- "ABB"
all_p + tdat_all_p + plot_layout(guides = 'collect', design = layout) & theme(legend.position = "bottom")
                   color = cluster,
                   group = cluster)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_pointrange(aes(xmin = ci.low, xmax = ci.up)) +
  scale_y_discrete(labels = Hugo_rownames) +
  scale_color_manual(values = cluster_colors) +
  theme_classic() +
  xlab("log odds ratio") +
  ylab("Gene") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y", switch = "y") +
  theme(
    panel.grid.major.y = element_line(colour = "#f0f0f0"),
    axis.title.y = element_blank()
  )

tdat_all <- all_muts %>%
  filter(Hugo_Symbol %in% all_compare_plot$Hugo_Symbol) %>%
  distinct(Tumor_Sample_Barcode, Hugo_Symbol, clust_knn_char, type) %>%
  select(type, clust_knn_char, Hugo_Symbol) %>%
  table()
tdat_all <- left_join(
  as.data.frame(tdat_all),
  clinical_plus_cluster %>%
    select(Tumor_Sample_Barcode, type, clust_knn_char) %>%
    distinct() %>%
    group_by(type, clust_knn_char) %>%
    summarise(ntumors = n())) %>%
  group_by(type, clust_knn_char) %>%
  mutate(c_prop = Freq / ntumors) %>%
  mutate(c_prop = ifelse(is.na(c_prop), 0, c_prop)) %>%
  select(type, clust_knn_char, Hugo_Symbol, c_prop) %>%
  rename(cluster = clust_knn_char)
tdat_all <- all_compare_plot %>% left_join(tdat_all)
tdat_all_p <- tdat_all %>% ggplot(aes(type, Row_Index, fill = c_prop)) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", "Proportion of \nTumor Type \nMutated Within Cluster", limits = c(0, 1)) +
  xlab("Tumor Type") +
  ylab("Gene Symbol") +
  scale_y_discrete(labels = Hugo_rownames) +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(colour = "#f0f0f0")
  )

layout <- "ABB"
all_p + tdat_all_p + plot_layout(guides = 'collect', design = layout) & theme(legend.position = "bottom")
```
